---
title: "CANOPE CNV Global Report"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: hide
params:
  cnv_calls: NA
  qc_metrics: NA
  counts: NA
  bed_file: NA
  models: NA
  refs: NA
  settings: NA
  sample_table: NA
  log_file: NA
  hmm_engine_used: NA
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)

library(DT)
library(plotly)
library(ggplot2)
library(dplyr)
library(knitr)
library(tidyr)
library(htmltools)
library(kableExtra)

# Metadata columns that are never sample columns in CANOPE's `counts` table
.canope_metadata_cols <- c("target", "chromosome", "start", "end", "gc", "GENE", "mean")

# ---------------------------------------------------------------------------
# Per-CNV plotting data, built the same way generate_plots.R does for the
# static PDFs, so the interactive report and the PDF plots always agree.
# Assumes the CANOPE package functions (targets_to_rows, parse_canope_targets,
# format_chr_label, compute_exon_index, model_lookup, model_matrix_lookup)
# are on the search path.
# ---------------------------------------------------------------------------
get_canope_plot_data <- function(call_row, counts, bed_file, models, refs, flank = 5) {

  sample <- call_row$SAMPLE
  target_bounds <- parse_canope_targets(call_row$TARGETS)
  if (length(target_bounds) < 2) return(NULL)
  idx_start <- target_bounds[1]
  idx_end   <- target_bounds[2]

  ref_samples <- refs[[sample]]
  if (is.null(ref_samples) || length(ref_samples) == 0) return(NULL)
  if (is.null(models[[sample]])) return(NULL)

  row_start <- targets_to_rows(idx_start, bed_file$target)
  row_end   <- targets_to_rows(idx_end, bed_file$target)
  if (is.na(row_start) || is.na(row_end)) return(NULL)

  exon_range <- seq(max(1, row_start - flank), min(nrow(bed_file), row_end + flank))

  target_chr <- format_chr_label(call_row$CHR)
  bed_chr    <- format_chr_label(bed_file$chromosome[exon_range])
  single_chr <- length(unique(bed_chr)) == 1
  if (!single_chr) exon_range <- exon_range[bed_chr == target_chr]
  if (length(exon_range) == 0) return(NULL)

  cnv_rows <- targets_to_rows(seq(idx_start, idx_end), bed_file$target)
  cnv_rows <- cnv_rows[!is.na(cnv_rows)]

  model_mean <- model_lookup(models[[sample]], "mean", bed_file$target[exon_range])
  model_var  <- model_lookup(models[[sample]], "var_estimate", bed_file$target[exon_range])

  pseudocount   <- 0.5
  # Raw counts, kept only for the display coverage panel (test_log / ref_cov),
  # matching the raw-coverage panel in the static PDFs.
  test_counts   <- counts[exon_range, sample]

  # The ratio/CI and z-score panels must use the model's own test_counts/
  # ref_matrix (not raw `counts`), since model_mean/model_var were fit on
  # call_cnvs()'s internally normalised data -- comparing raw counts against
  # that normalised model would mix two depth scales.
  test_counts_model <- model_lookup(models[[sample]], "test_counts", bed_file$target[exon_range])
  expected_safe <- pmax(model_mean, 1) + pseudocount

  ratio      <- log2((test_counts_model + pseudocount) / expected_safe)
  dispersion <- model_var - model_mean
  size_param <- ifelse(dispersion > 0, model_mean^2 / dispersion, Inf)
  alpha <- 0.05
  mins <- stats::qnbinom(alpha / 2,     mu = model_mean, size = size_param)
  maxs <- stats::qnbinom(1 - alpha / 2, mu = model_mean, size = size_param)
  lo <- log2((mins + pseudocount) / expected_safe)
  hi <- log2((maxs + pseudocount) / expected_safe)

  exon_labels <- compute_exon_index(bed_file)[exon_range]

  # Z-scores vs. references. Uses the same robust, floored var_estimate as
  # the ratio/NB-interval panel above (rather than a raw per-target SD across
  # just the handful of references), so all panels agree with the model and
  # with each other.
  ref_mat <- model_matrix_lookup(models[[sample]], "ref_matrix", bed_file$target[exon_range])
  z_means <- rowMeans(ref_mat, na.rm = TRUE)
  z_sds   <- sqrt(pmax(model_var, 1))
  z_sds[!is.finite(z_sds) | z_sds < 1e-6] <- 1e-6
  sample_z <- (test_counts_model - z_means) / z_sds
  ref_z    <- sapply(seq_along(ref_samples), function(i) (ref_mat[, i] - z_means) / z_sds)

  ref_cov <- do.call(rbind, lapply(ref_samples, function(r) {
    data.frame(
      Exon = exon_range, ExonLabel = exon_labels,
      Coverage = log2(counts[exon_range, r] + pseudocount),
      Group = "Reference samples", RefName = r,
      stringsAsFactors = FALSE
    )
  }))

  # Per-call diagnostic: does this window's background (non-called)
  # coverage fit the model as well as the interval implies?
  bg_calib <- check_background_calibration(ratio, lo, hi, exon_range %in% cnv_rows)

  list(
    sample      = sample,
    exon_range  = exon_range,
    exon_labels = exon_labels,
    affected    = exon_range %in% cnv_rows,
    test_log    = log2(test_counts + pseudocount),
    bg_calibration = bg_calib,
    ratio_df    = data.frame(
      Exon = exon_range, ExonLabel = exon_labels,
      Ratio = ratio, CI_low = lo, CI_high = hi,
      Affected = exon_range %in% cnv_rows
    ),
    z_df = data.frame(
      Exon = exon_range, ExonLabel = exon_labels,
      Z = sample_z, Affected = exon_range %in% cnv_rows
    ),
    ref_z_df = {
      rz <- as.data.frame(ref_z)
      colnames(rz) <- ref_samples
      rz$Exon <- exon_range
      rz$ExonLabel <- exon_labels
      tidyr::pivot_longer(rz, cols = all_of(ref_samples), names_to = "RefName", values_to = "Z")
    },
    ref_cov     = ref_cov,
    ref_samples = ref_samples
  )
}
```

<style>
  .alert-warning { border-left: 5px solid #f0ad4e; margin-bottom: 20px; background-color: #fcf8e3; color: #8a6d3b; padding: 15px; border-radius: 4px; }
  .dt-buttons { margin-bottom: 15px; }
  .btn-default { background-color: #34495e !important; color: white !important; border: none !important; }
  .btn-default:hover { background-color: #2c3e50 !important; }
</style>

## Summary

- **Number of samples with CNV calls**:
  `r length(unique(params$cnv_calls$SAMPLE))`

- **Total CNV calls**:
  `r nrow(params$cnv_calls)`

- **HMM engine**:
  `r if (identical(params$hmm_engine_used, "legacy_canoes")) "Legacy CANOES (original published HMM core)" else "CANOPE (modern engine)"`

- **Report generated**:
  `r Sys.time()`

## QC Metrics Overview (all samples)

```{r qc_overview, results='asis'}
if (!is.null(params$qc_metrics) && is.data.frame(params$qc_metrics) && nrow(params$qc_metrics) > 0) {

  missing_chrs <- params$qc_metrics[params$qc_metrics$Type == "Missing Chromosome", ]
  if (nrow(missing_chrs) > 0) {
    for (i in seq_len(nrow(missing_chrs))) {
      # Each cat() call must start at column 0: results='asis' output goes
      # straight into the Markdown source, and Pandoc treats an indented
      # line (4+ spaces) as a code block rather than raw HTML.
      cat('<div class="alert alert-warning" role="alert">\n')
      cat('<h4 class="alert-heading" style="margin-top:0;font-weight:bold;"><i class="fa fa-exclamation-triangle"></i> Pipeline Warning</h4>\n')
      cat(sprintf('<p style="margin-bottom:0;">%s. <strong>Analysis for this chromosome has been skipped automatically.</strong></p>\n',
                  htmltools::htmlEscape(missing_chrs$Details[i])))
      cat('</div>\n')
    }
    cat("<br>")
  }

  display_qc <- params$qc_metrics[params$qc_metrics$Type != "Missing Chromosome", ]

  if (nrow(display_qc) > 0) {
    DT::datatable(
      display_qc,
      extensions = 'Buttons',
      options = list(
        pageLength = 15,
        scrollX = TRUE,
        dom = 'Bfrtip',
        buttons = list(
          list(extend = 'colvis', text = '<i class="fa fa-eye"></i> Show/Hide Columns',
               collectionLayout = 'fixed columns', fade = 100),
          'copy', 'csv'
        )
      ),
      caption = "QC flags across all samples"
    )
  } else {
    cat("<p>No sample-specific QC warnings or issues recorded.</p>\n")
  }
} else {
  cat("<p><em>No QC metrics file supplied — QC section skipped.</em></p>\n")
}
```

---

## Principal Component Analysis (PCA) of Coverage Profiles

```{r pca_coverage}
counts <- params$counts
if (!is.null(counts) && is.data.frame(counts) && ncol(counts) >= 3) {

  sample_cols <- setdiff(colnames(counts), .canope_metadata_cols)

  if (length(sample_cols) >= 3) {
    count_mat <- as.matrix(counts[, sample_cols, drop = FALSE])
    log_mat <- log2(count_mat + 1)
    row_var <- apply(log_mat, 1, var, na.rm = TRUE)
    keep <- row_var > 0 & !is.na(row_var)

    if (sum(keep) >= 2) {
      log_mat <- log_mat[keep, ]
      pca <- prcomp(t(log_mat), scale. = TRUE, center = TRUE)
      var_exp <- summary(pca)$importance[2, ] * 100
      pca_df <- data.frame(PC1 = pca$x[, 1], PC2 = pca$x[, 2], Sample = sample_cols)

      sample_table <- params$sample_table
      if (!is.null(sample_table) && is.data.frame(sample_table) &&
          "gender" %in% colnames(sample_table) && "sample_name" %in% colnames(sample_table)) {
        pca_df <- merge(pca_df, sample_table[, c("sample_name", "gender")],
                        by.x = "Sample", by.y = "sample_name", all.x = TRUE)
        color_by <- "gender"
      } else {
        pca_df$Group <- "All samples"
        color_by <- "Group"
      }

      p <- plot_ly(pca_df, x = ~PC1, y = ~PC2, text = ~Sample,
                   color = as.formula(paste0("~", color_by)),
                   type = 'scatter', mode = 'markers+text',
                   textposition = 'top right',
                   marker = list(size = 10),
                   hovertemplate = paste('<b>%{text}</b><br>PC1: %{x:.2f}<br>PC2: %{y:.2f}<extra></extra>')) %>%
        layout(title = paste("PCA of Coverage Profiles (", ncol(count_mat), " samples, ", nrow(log_mat), " targets)"),
               xaxis = list(title = paste0("PC1 (", round(var_exp[1], 1), "%)")),
               yaxis = list(title = paste0("PC2 (", round(var_exp[2], 1), "%)")),
               legend = list(title = list(text = ifelse(color_by == "gender", "Gender", "Group"))))
      p
    } else {
      cat("Insufficient variation for PCA after filtering (less than 2 variable targets).")
    }
  } else {
    cat("Not enough samples (≥3 required) for PCA.")
  }
} else {
  cat("Counts data not available or insufficient dimensions for PCA.")
}
```

# Per-sample Reports

```{r per_sample_loop, results='asis'}
samples <- unique(params$cnv_calls$SAMPLE)

if (length(samples) == 0) {

  cat("No CNV calls found in any sample.\n")

} else {

  for (s in samples) {

    cat("\n\n## ", s, "\n")

    # QC metrics for this sample
    if (!is.null(params$qc_metrics) && is.data.frame(params$qc_metrics)) {
      qc_s <- params$qc_metrics[
        (params$qc_metrics$Sample == s | params$qc_metrics$Sample == "All") &
         params$qc_metrics$Type != "Missing Chromosome",
      ]
      if (nrow(qc_s) > 0) {
        cat("\n### QC flags\n")
        w_qc <- DT::datatable(qc_s, options = list(dom = 't', pageLength = 5))
        rendered_qc <- htmltools::renderTags(w_qc)
        knitr::knit_meta_add(rendered_qc$dependencies)
        cat(rendered_qc$html)
        cat("\n")
      }
    }

    # Key metrics for this sample
    if (!is.null(params$counts) && s %in% colnames(params$counts)) {
      cov <- params$counts[, s]
      median_depth <- median(cov, na.rm = TRUE)
      total_reads  <- sum(cov, na.rm = TRUE)

      cat("\n### Key metrics\n")
      cat(paste0("- Median target depth: ", round(median_depth, 1), "\n"))
      cat(paste0("- Total reads: ", format(total_reads, big.mark = ","), "\n"))

      if (!is.null(params$settings$qc_min_cov) && median_depth < params$settings$qc_min_cov) {
        cat("⚠️ **Below minimum depth threshold**\n")
      }
      if (!is.null(params$settings$qc_min_total_reads) && total_reads < params$settings$qc_min_total_reads) {
        cat("⚠️ **Below minimum total reads**\n")
      }
    }

    # CNV table for this sample
    cnv_s <- params$cnv_calls[params$cnv_calls$SAMPLE == s, ]

    cat("\n### CNV calls\n")

    if (nrow(cnv_s) > 0) {

      display_cols <- intersect(
        c("SAMPLE", "CHR", "GENE", "INTERVAL", "KB", "CNV", "NUM_TARG",
          "CN_label", "Q_SOME", "NUM_REFS", "Confidence"),
        colnames(cnv_s)
      )
      cnv_display <- cnv_s[, display_cols, drop = FALSE]
      cnv_display$KB <- round(as.numeric(cnv_display$KB), 2)
      rownames(cnv_display) <- NULL

      dt_opts <- list(
        dom = "Bfrtip",
        buttons = list(
          list(extend = 'colvis', text = '<i class="fa fa-eye"></i> Show/Hide Columns',
               collectionLayout = 'fixed columns', fade = 100),
          "copy", "csv", "excel"
        ),
        autoWidth = FALSE
      )

      w_cnv <- DT::datatable(
        cnv_display, rownames = FALSE,
        extensions = "Buttons", options = dt_opts,
        caption = paste("Detected CNVs for", s),
        width = "100%"
      )
      if ("Confidence" %in% colnames(cnv_display)) {
        w_cnv <- w_cnv %>%
          formatStyle("Confidence",
                     backgroundColor = styleEqual(
                       c("HIGH", "MEDIUM", "LOW"),
                       c("lightgreen", "yellow", "salmon")
                     ))
      }

      rendered_cnv <- htmltools::renderTags(w_cnv)
      knitr::knit_meta_add(rendered_cnv$dependencies)
      cat(rendered_cnv$html)
      cat("\n")

    } else {
      cat("No CNV calls detected for this sample.\n")
    }

    # Warning for low-confidence genes
    low_conf_genes <- params$settings$low_confidence_genes
    all_genes <- unique(trimws(unlist(strsplit(as.character(cnv_s$GENE), "[;, ]"))))
    flag <- intersect(all_genes, low_conf_genes)
    if (length(flag) > 0) {
      cat("\n⚠️ **Warning:** This sample has CNVs involving genes with known pseudogene/homology issues: ",
          paste(flag, collapse = ", "), ". Consider orthogonal validation.\n")
    }

    # Interactive plots for each CNV
    if (nrow(cnv_s) > 0) {

      cat("\n### CNV plots\n")

      for (i in seq_len(nrow(cnv_s))) {

        call <- cnv_s[i, ]

        cat("\n#### ", call$GENE, " (", call$CNV, ")\n", sep = "")
        conf_str <- if ("Confidence" %in% colnames(call)) call$Confidence else "n/a"
        cat("*Confidence: ", conf_str, " | Q_SOME: ", call$Q_SOME,
            " | NUM_REFS: ", call$NUM_REFS, "*\n\n", sep = "")

        plot_data <- get_canope_plot_data(call, params$counts, params$bed_file,
                                          params$models, params$refs)

        if (is.null(plot_data)) {
          cat("*(Insufficient data to generate plot)*\n")
          next
        }

        # Per-call diagnostic: flags when the fraction of non-called exons in
        # this window that fall outside the modelled 95% interval is
        # statistically higher than the ~5% expected under a well-calibrated
        # interval. Informational only -- doesn't change the call, interval,
        # or confidence score. See check_background_calibration() in
        # canope_utils.R.
        if (isTRUE(plot_data$bg_calibration$flag)) {
          cat("\n> \u26a0\ufe0f **Background calibration flag:** ",
              round(plot_data$bg_calibration$pct_outside), "% of non-called exons in this window (",
              plot_data$bg_calibration$n_outside, "/", plot_data$bg_calibration$n_background,
              ") fall outside the modeled 95% interval \u2014 well above the ~5% expected under a ",
              "well-calibrated model. This is a diagnostic flag, not a correction: it can mean the real ",
              "alteration extends beyond the exons called here, the reference set is a weaker match for ",
              "this sample/region, or there is a technical/batch difference at this locus. Worth a manual ",
              "look before reporting this call with full confidence.\n\n", sep = "")
        }

        ci_subtitle <- "95% NB predictive interval"
        if (isTRUE(plot_data$bg_calibration$flag)) {
          ci_subtitle <- paste0("95% NB predictive interval  [", round(plot_data$bg_calibration$pct_outside),
                                "% of background exons outside CI (", plot_data$bg_calibration$n_outside, "/",
                                plot_data$bg_calibration$n_background, ") \u2014 check region/reference match]")
        }

        # --- Panel 1: log2(Observed / Expected) ratio with NB CI ribbon ---
        p_ratio <- ggplot(plot_data$ratio_df, aes(x = Exon)) +
          geom_ribbon(aes(ymin = CI_low, ymax = CI_high), fill = "grey85", alpha = 0.5) +
          geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
          geom_point(data = subset(plot_data$ratio_df, !Affected),
                    aes(y = Ratio, color = "Reference range",
                        text = paste0("Sample: ", plot_data$sample, "<br>Exon: ", ExonLabel,
                                      "<br>log2(Obs/Exp): ", round(Ratio, 3))),
                    size = 3) +
          geom_point(data = subset(plot_data$ratio_df, Affected),
                    aes(y = Ratio, color = "Affected exon(s)",
                        text = paste0("Sample: ", plot_data$sample, "<br>Exon: ", ExonLabel,
                                      "<br>log2(Obs/Exp): ", round(Ratio, 3),
                                      "<br>Confidence: ", conf_str)),
                    size = 3.5) +
          labs(y = "log2(Observed / Expected)", x = NULL, subtitle = ci_subtitle) +
          theme_bw() +
          scale_color_manual(values = c("Reference range" = "steelblue", "Affected exon(s)" = "red"), name = NULL) +
          theme(legend.position = "top", axis.text.x = element_text(angle = 45, size = 8, hjust = 1)) +
          scale_x_continuous(breaks = plot_data$ratio_df$Exon, labels = plot_data$ratio_df$ExonLabel)

        # --- Panel 2: log2 coverage (test + references) ---
        obs_df <- data.frame(
          Exon = plot_data$exon_range, ExonLabel = plot_data$exon_labels,
          Coverage = plot_data$test_log,
          Group = ifelse(plot_data$affected, "Affected exon(s)", "Test sample"),
          SampleID = plot_data$sample,
          stringsAsFactors = FALSE
        )
        p_cov_data <- bind_rows(plot_data$ref_cov, obs_df)

        p_cov <- ggplot() +
          geom_point(data = subset(p_cov_data, Group == "Reference samples"),
                    aes(x = Exon, y = Coverage, color = Group,
                        text = paste0("Sample: ", RefName, "<br>Exon: ", ExonLabel,
                                      "<br>log2 coverage: ", round(Coverage, 3))),
                    alpha = 0.6, size = 2.5) +
          geom_point(data = subset(p_cov_data, Group == "Test sample"),
                    aes(x = Exon, y = Coverage, color = Group,
                        text = paste0("Sample: ", SampleID, "<br>Exon: ", ExonLabel,
                                      "<br>log2 coverage: ", round(Coverage, 3))),
                    size = 2.5) +
          geom_point(data = subset(p_cov_data, Group == "Affected exon(s)"),
                    aes(x = Exon, y = Coverage, color = Group,
                        text = paste0("Sample: ", SampleID, "<br>Exon: ", ExonLabel,
                                      "<br>log2 coverage: ", round(Coverage, 3),
                                      "<br>Confidence: ", conf_str)),
                    size = 3.5) +
          labs(y = "log2(Coverage + 0.5)", x = NULL) +
          theme_bw() +
          scale_color_manual(values = c("Reference samples" = "grey50", "Test sample" = "blue",
                                        "Affected exon(s)" = "red"), name = NULL) +
          theme(legend.position = "top", axis.text.x = element_text(angle = 45, size = 8, hjust = 1)) +
          scale_x_continuous(breaks = plot_data$exon_range, labels = plot_data$exon_labels)

        # --- Panel 3 (NEW): Z-score vs. references ---
        z_ref_df <- plot_data$ref_z_df
        z_ref_df$IsTest <- FALSE
        z_test_df <- plot_data$z_df
        z_test_df$RefName <- s
        z_test_df$IsTest <- TRUE

        p_z <- ggplot() +
          geom_line(data = z_ref_df, aes(x = Exon, y = Z, group = RefName),
                   colour = "gray60", linewidth = 0.6, alpha = 0.7) +
          geom_point(data = z_ref_df,
                    aes(x = Exon, y = Z, group = RefName,
                        text = paste0("Sample: ", RefName, "<br>Exon: ", ExonLabel,
                                      "<br>Z-score: ", round(Z, 2))),
                    colour = "gray60", size = 1.5, alpha = 0.7) +
          geom_line(data = z_test_df, aes(x = Exon, y = Z, group = RefName),
                   colour = "red", linewidth = 1.1) +
          geom_point(data = subset(z_test_df, !Affected),
                    aes(x = Exon, y = Z,
                        text = paste0("Sample: ", RefName, "<br>Exon: ", ExonLabel,
                                      "<br>Z-score: ", round(Z, 2))),
                    colour = "red", size = 2.2) +
          geom_point(data = subset(z_test_df, Affected),
                    aes(x = Exon, y = Z,
                        text = paste0("Sample: ", RefName, "<br>Exon: ", ExonLabel,
                                      "<br>Z-score: ", round(Z, 2),
                                      "<br>Confidence: ", conf_str)),
                    colour = "darkred", size = 2.8) +
          geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
          labs(y = "Z-score vs. references", x = NULL,
               subtitle = "Test sample (red) — gray lines are individual reference samples") +
          theme_bw() +
          theme(legend.position = "none", axis.text.x = element_text(angle = 45, size = 8, hjust = 1)) +
          scale_x_continuous(breaks = plot_data$exon_range, labels = plot_data$exon_labels)

        w_ratio <- ggplotly(p_ratio, tooltip = "text", height = 360) %>%
          layout(margin = list(t = 60, b = 80), legend = list(orientation = "h", y = 1.15))
        w_cov <- ggplotly(p_cov, tooltip = "text", height = 360) %>%
          layout(margin = list(t = 60, b = 80), legend = list(orientation = "h", y = 1.15))
        w_z <- ggplotly(p_z, tooltip = "text", height = 360) %>%
          layout(margin = list(t = 60, b = 80))

        for (w in list(w_ratio, w_cov, w_z)) {
          rendered <- htmltools::renderTags(w)
          knitr::knit_meta_add(rendered$dependencies)
          cat(rendered$html)
          cat("\n")
        }
      }
    }
  }
}
```

---
## Confidence Score Method

<details>
<summary><strong>&#9432; Click to expand/collapse confidence scoring rules</strong></summary>

```{r confidence_recap, results='asis'}
settings <- params$settings

score_matrix <- data.frame(
  Category = c("**HIGH**", "**MEDIUM**", "**LOW**"),
  `Q_SOME (phred)` = c(
    sprintf("≥ %s", settings$high_q_score %||% 50),
    sprintf("≥ %s", settings$med_q_score %||% 20),
    "Any"
  ),
  `Reference samples` = c(
    sprintf("≥ %s", settings$high_num_refs %||% 10),
    sprintf("≥ %s", settings$med_num_refs %||% 5),
    "Any"
  ),
  `MLCN consistency` = c(
    "Consistent with call (DEL≤1, DUP≥3)",
    "Not required",
    "Any"
  ),
  check.names = FALSE,
  stringsAsFactors = FALSE
)

forced_genes <- settings$low_confidence_genes
if (is.null(forced_genes)) forced_genes <- c()

cat('<div style="margin: 10px 0 20px 0;">\n')
cat('<p><strong>Assignment rules:</strong></p>\n')
cat('<ul>\n')
cat('<li><strong>HIGH</strong> only if <em>all</em> HIGH criteria are met.</li>\n')
cat('<li><strong>MEDIUM</strong> if HIGH not met but <em>all</em> MEDIUM criteria are met.</li>\n')
cat('<li><strong>LOW</strong> otherwise.</li>\n')
if (length(forced_genes) > 0) {
  cat(sprintf('<li><strong>Always LOW</strong> (genes with homology issues): <em>%s</em>.</li>\n',
              paste(forced_genes, collapse = ", ")))
}
cat('</ul>\n')
cat('</div>\n')

knitr::kable(score_matrix, format = "html", escape = FALSE, align = c("l", "c", "c", "l")) %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                            full_width = FALSE, position = "left")
```

</details>

## Pipeline Log

```{r log_warnings, results='asis'}
if (!is.null(params$log_file) && !identical(params$log_file, NA) && file.exists(params$log_file)) {
  log_lines <- readLines(params$log_file, warn = FALSE)

  warning_lines <- log_lines[startsWith(log_lines, "[WARNING]")]
  session_lines <- log_lines[grepl("^(R version|Platform|  )", log_lines)]
  info_lines    <- log_lines[!startsWith(log_lines, "[WARNING]") &
                              !grepl("^(R version|Platform|  )", log_lines)]

  cat('<details>\n<summary><strong>&#9888; Warnings (', length(warning_lines), ')</strong></summary>\n\n', sep = "")
  if (length(warning_lines) > 0) {
    cat('<pre style="max-height:400px;overflow-y:auto;background:#fff8e1;border-left:4px solid #f9a825;padding:8px;">\n')
    cat(paste(htmltools::htmlEscape(warning_lines), collapse = "\n"))
    cat('\n</pre>\n')
  } else {
    cat('<p><em>No warnings recorded.</em></p>\n')
  }
  cat('</details>\n\n')

  cat('<details>\n<summary><strong>&#128196; Full pipeline log</strong></summary>\n\n')
  cat('<pre style="max-height:500px;overflow-y:auto;background:#f5f5f5;padding:8px;">\n')
  cat(paste(htmltools::htmlEscape(info_lines), collapse = "\n"))
  cat('\n</pre>\n')
  cat('</details>\n\n')

} else {
  cat("*Log file not available.*\n")
}
```

## Session Info

```{r session_info, results='asis'}
si <- sessionInfo()
si_lines <- capture.output(print(si))

cat('<details>\n<summary><strong>&#128196; Session information</strong></summary>\n\n')
cat('<pre style="max-height:400px;overflow-y:auto;background:#f5f5f5;padding:8px;">\n')
cat(paste(htmltools::htmlEscape(si_lines), collapse = "\n"))
cat('\n</pre>\n')
cat('</details>\n')
```
